---
title: "Intro to Spatial Data in R"
author: "Brooke Helmick"
format: html
---

## Downloading data

```{r}
knb_url <- "https://dev.nceas.ucsb.edu/knb/d1/mn/v2/object/urn%3Auuid%3Aaceaecb2-1ce0-4d41-a839-d3607d32bb58"

download.file(url = knb_url, destfile = '../shapefile_demo_data.zip')

unzip('../shapefile_demo_data.zip', exdir = '../data')

file.remove('../shapefile_demo_data.zip')
```

## Set Up 

```{r}
#| message: false

library(readr)
library(sf)
library(ggplot2)
library(leaflet)
library(scales)
library(ggspatial)
library(dplyr)
library(janitor)
library(here)
```

## Read region data 

```{r}
ak_regions <- read_sf(here("data", "ak_regions_simp.shp"))

plot(ak_regions)

class(ak_regions)

head(ak_regions)

colnames(ak_regions)

glimpse(ak_regions)

st_crs(ak_regions)
```

## Transforming CRS

```{r}
ak_regions_3338 <- ak_regions %>%
  st_transform(crs = 3338)

plot(ak_regions_3338)
```

## 'sf' and the 'tidyverse' 

```{r}
ak_regions_3338 %>% 
  select(region)

## geometry column is a sticky column, will always be there
```

## Filter by region 

```{r}
ak_regions_3338 %>% 
  select(region)
```

## Filter for a specific region 

```{r}
ak_regions_3338 %>%
    filter(region == "Southeast")
```

## Read the population data 

```{r}
pop <- read_csv(here("data", "alaska_population.csv"))
head(pop)
class(pop)
```

## Convert pop data into a spatial object 

```{r}
pop_4326 <- st_as_sf(pop, 
                     coords = c("lng", "lat"), 
                     crs = 4326, 
                     remove = F)
```


## Joining population data with region data 

```{r}
pop_3338 <- st_transform(pop_4326, 
                         crs = 3338)

pop_joined <- st_join(pop_3338, 
                      ak_regions_3338, 
                      join = st_within)

head(pop_joined)
```

## Calculating total population by region 

```{r}
pop_region <- pop_joined %>% 
  as.data.frame() %>% 
  group_by(region) %>% 
  summarize(total_pop = sum(population))

class(pop_region)
head(pop_region)
```

## Joining Regions with total population by region 

```{r}
pop_region_3338 <- left_join(ak_regions_3338, 
                             pop_region, 
                             by = "region")
plot(pop_region_3338["total_pop"])
```

## Calculate total population by management area 

```{r}
pop_mgmt_3338 <- pop_region_3338 %>% 
  group_by(mgmt_area) %>% 
  summarize(total_pop_mgmt = sum(total_pop), 
            do_union = F)

plot(pop_mgmt_3338["total_pop_mgmt"])
```

## Visualize with ggplot 

```{r}
ggplot(pop_region_3338) + 
  geom_sf(aes(fill = total_pop)) + 
  labs(fill = "Total Population") + 
  scale_fill_continuous(low = "khaki", 
                        high = "firebrick", 
                        labels = comma) + 
  theme_bw()
```

