---
title: "Quick Analysis"
author: "Brooke Helmick"
format: html
editor: visual
execute: 
  warning: false 
  message: false 
---

# Introduction

## About the data

The data for this analysis was downloaded from the Artctic Data Center on 10/7/2024: [Arctic Data Center](https://arcticdata.io/catalog/view/doi:10.18739/A25T3FZ8X)

## Set up

Attach important packages. 

```{r}
library(readr)
library(here)
```

## Read in the data

```{r}
#| message: false 
#| warning: false
bg_chem <- read_csv(here('data', 'BGchem2008data.csv'))
```

# Analysis

## Calculate summary statistics

```{r}
#| eval: false 
#| include: false 

# print the columnnames 
colnames(bg_chem)

# get the general structure 
str(bg_chem)

# First six lines 
head(bg_chem)

# get a summary of all columns 
summary(bg_chem)

# unique values of a column 
unique(bg_chem$date)

```

Calculate the summary statistics (means) for nitrate, nitrite, ammonia, and phosphate. 

```{r}
nitrate <- mean(bg_chem$NO3)
nitrite <- mean(bg_chem$NO2)
amm <- mean(bg_chem$NH4)
phos <- mean(bg_chem$P)
```

## Calculate the mean Redfield ratio

Ratio of nitrogen (nitrate, nitrite, amm) to phos, i.e. nutrient availability in a body of water

```{r}
ratio2 <- (nitrate + nitrite + amm) / phos
```

The Redfield ratio for this data set is approximately `r round(ratio2)`

## Plot Redfield ratio

```{r}
plot(bg_chem$P, bg_chem$NO2 + bg_chem$NO3 + bg_chem$NH4)
```


# Conclusion
