---
title: "Intro to Quarto"
author: "Brooke Helmick"
format: html
editor: visual
execute: 
  warning: false
  message: false 
bibliography: references.bib
---

# Level 1 header

this is regular plain old boring text

## Level 2 header

-   blank line between sections

-   doesn't always make a difference but sometimes it does

-   bullet item 2

-   3 etc

-   item 4 list

    -   used tab in visual editor

1.  numbered list
2.  item 2

here is some more plain text. But if I want *italics* or *italics* \*\* or \_x \_

**bold** two asterisks or \_\_bold \_\_

***bold = three italics***

back to normal text – important to note that you cannot use syntax in visual editor

Create a code chunk:

```{r important r code}

```

| Col1    | Col2 | Col3 |
|---------|------|------|
| thing 1 | 1    |      |
| thing 2 | 2    |      |
| thing 3 | 3    |      |
| thing 4 | 4    |      |
| thing 5 | 5    |      |

```{x <- 1 + 2}
```

table 1

here is a citation [@oke2020] – runs like zotero bibtech is a language for citation management

```{r important code 2}
x <- 1 + 2
```

My important result is `r x` (this is an in-line code chunk_ or another important thing is `r 6 * 7`)

if you want text to look like code: `this is code formatted`

```{r}

x <- 4 * 8 

height_ft <- c(5.2, 6.0, 5.7)

coef <- 3.14
```

